A Wait-free Queue with Polylogarithmic Step Complexity

### 2022-12-29 podc_paper/algorithm.tex
working on code

\begin{figure}
\begin{minipage}[t]{0.465\textwidth}
\begin{algorithmic}[1]
\setcounter{ALG@line}{0}

@@

\end{algorithmic}
\end{minipage}
\begin{minipage}[t]{0.525\textwidth}

@@

\Function{Block}{CreateBlock}{\var{Node v, int i}} 
    \State\linecomment{create new block for a \op{Refresh} to install in \var{v.blocks[i]}}
    \State\linecomment{(return \var{null} if no blocks need to be propagated to \var{v})}
    \State let \var{new} be a new \var{Block} \label{initNewBlock}
    \State \var{new.\eleft} \assign\ $\var{v.left.head} - 1$
    \State \var{new.\eright} \assign\ $\var{v.right.head} - 1$
	\State \hangbox{\var{new.sum\sub{enq}} \assign\ \var{v.left.blocks[new.\eleft].sum\sub{enq} + \\
			v.right.blocks[new.\eright].sum\sub{enq}}}
	\State \hangbox{\var{new.sum\sub{deq}} \assign\ \var{v.left.blocks[new.\eleft].sum\sub{deq} + \\
			v.right.blocks[new.\eright].sum\sub{deq}}}
    \State \var{num\sub{enq}} \assign\ \var{new.sum\sub{enq} - v.blocks[i-1].sum\sub{enq}}
    \State \var{num\sub{deq}} \assign\ \var{new.sum\sub{deq} - v.blocks[i-1].sum\sub{deq}}
    \If{$\var{v} = \var{root}$}
        \State \hangbox{\var{new.size \assign\ max(0, v.blocks[i-1].\size\ + num\sub{enq} - num\sub{deq})}\label{computeLength}}
    \EndIf
    \If{$\var{num\sub{enq}} + \var{num\sub{deq}} = 0$}
        \State \Return \var{null}
    \Else
        \State \Return \var{new}
    \EndIf
\EndFunction{CreateBlock}

@@

\Function{<int, int>}{IndexDequeue}{\var{Node v, int b, int i}}
    \State \linecomment{return \var{<x, y>} such that \var{i}th dequeue in D(\var{v.blocks[b]})}
    \State \linecomment{is \var{y}th dequeue of D(\var{root.blocks[x]})}
    \State \linecomment{Precondition: \var{v.blocks[b]} has been propagated to root}
    \State \linecomment{and $\var{v.blocks[b].num\sub{deq}}\geq \var{i}$}
    \If{$\var{v} = \var{root}$}
        \State\Return \var{<b, i>} \label{indexBaseCase}
    \Else
	    \State \var{dir \assign\ (v.parent.left = v ? left : right)} 
    	\State \var{sup \assign\ v.blocks[b].super}
	    \If{$\var{b} > \var{v.parent.blocks[sup].end\sub{dir}}$}
	        \State{\var{sup} \assign\ $\var{sup}+1$}
	    \EndIf
	    \State \hangbox{\var{i += v.blocks[b-1].sum\sub{deq} - \\
	    		v.blocks[v.parent.blocks[sup-1].end\sub{dir}].sum\sub{deq}}}
        \If{$\var{dir} = \var{right}$} 
        	\State \hangbox{\var{i += v.blocks[v.parent.blocks[sup].end\sub{left}].sum\sub{deq} -\\
					v.blocks[v.parent.blocks[sup-1].end\sub{left}].sum\sub{deq}}\label{considerLeftBeforeRight}}
        \EndIf \label{computeISuperEnd}
        \State \Return\Call{IndexDequeue}{v.parent, sup, i}
    \EndIf
\EndFunction{IndexDequeue}

@@

\Function{element}{FindResponse}{\var{int b, int i}}
    \State \linecomment{find response to \var{i}th dequeue in \var{root.blocks[b]}}
    \If{\var{root.blocks[b-1].\size + root.blocks[b].num\sub{enq}} $<$ \var{i}}\label{checkEmpty}
        \State \linecomment{queue is empty when the dequeue is linearized}
        \State \Return \var{null} \label{returnNull}
    \Else \ \linecomment{response is the \var{e}th enqueue in the root}
        \State \hangbox{\var{e} \assign\ \var{i + root.blocks[b-1].sum\sub{enq} - \\
	root.blocks[b-1].size}\label{computeE}}
        \State \Return \var{root.GetEnqueue(root.\Call{DoublingSearch}{e, b})}\label{findAnswer}
    \EndIf
\EndFunction{FindResponse}

@@

\Function{<int, int>}{DoublingSearch}{\sl{int} e, \sl{int} end}
\State \tt{start \assign\ end-1} \label{dsearchStart}
\While{\tt{root.blocks[start].sum\sub{enq}}$> \assign\ $\tt{e}}
\State \tt{start \assign\ max(start-(end-start), 0)} \label{doubling}
\EndWhile \label{dsearchEnd}
\State \tt{b \assign\ root.BinarySearch(e, start, end)} \label{dsearchBinarySearch}
\State \tt{i \assign\ e- root.blocks[b-1].sum\sub{enq}} \label{DSearchComputei}
\State\Return \tt{<b,i>}
\EndFunction{DoublingSearch}
\end{algorithmic}
\end{algorithm}

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{blocks[b].num\sub{enq}$\geq$i$\geq 1$}}
\Function{element}{GetEnqueue}{\sl{int} b, \sl{int} i} \cmt{Returns the \tt{element} of $E_\tt{i}(\tt{this},\tt{b})$.}
\If{\tt{this} \bf{is} \tt{leaf}}
\State\Return \tt{blocks[b].element} \label{getBaseCase}
\ElsIf{\tt{i <= blocks[b].num\sub{enq-left}}} \label{leftOrRight} \cmt{$E_\tt{i}(\tt{this},\tt{b})$ is in the left child of this node.}
\State \tt{subblockIndex \assign left.BinarySearch(i+blocks[b-1].sum\sub{enq-left}, blocks[b-1].\eleft+1,} \label{leftChildGet}
\Statex \hspace{10.7em}\tt{blocks[b].\eleft)} 
\State \Return\tt{left.}\Call{GetEnqueue}{subblockIndex, i} 
\Else
\State \tt{i \assign\ i-blocks[b].num\sub{enq-left}}
\State\tt{subblockIndex \assign\ right.BinarySearch(i+blocks[b-1].sum\sub{enq-right}, blocks[b-1].\eright+1,} \label{rightChildGet}
\Statex \hspace{11.1em}\tt{blocks[b].\eright)} 
\State \Return\tt{right.}\Call{GetEnqueue}{subblockIndex, i} 
\EndIf
\EndFunction{GetEnqueue}

>>>>

\begin{figure}
\begin{minipage}[t]{0.461\textwidth}
\begin{algorithmic}[1]
\setcounter{ALG@line}{0}

@@

\end{algorithmic}
\end{minipage}
\begin{minipage}[t]{0.533\textwidth}

@@

\Function{Block}{CreateBlock}{\var{Node v, int i}} 
    \State\linecomment{create new block for a \op{Refresh} to install in \var{v.blocks[i]}}
    \State\linecomment{(return \var{null} if no blocks need to be propagated to \var{v})}
    \State let \var{new} be a new \var{Block} \label{initNewBlock}
    \State \var{new.\eleft} \assign\ $\var{v.left.head} - 1$
    \State \var{new.\eright} \assign\ $\var{v.right.head} - 1$
	\State \hangbox{\var{new.sum\sub{enq}} \assign\ \var{v.left.blocks[new.\eleft].sum\sub{enq}} + \\
			\var{v.right.blocks[new.\eright].sum\sub{enq}}}
	\State \hangbox{\var{new.sum\sub{deq}} \assign\ \var{v.left.blocks[new.\eleft].sum\sub{deq}} + \\
			\var{v.right.blocks[new.\eright].sum\sub{deq}}}
    \State \var{num\sub{enq}} \assign\ \var{new.sum\sub{enq} - v.blocks[i-1].sum\sub{enq}}
    \State \var{num\sub{deq}} \assign\ \var{new.sum\sub{deq} - v.blocks[i-1].sum\sub{deq}}
    \If{$\var{v} = \var{root}$}
        \State \hangbox{\var{new.size \assign\ max(0, v.blocks[i-1].\size\ + num\sub{enq} - num\sub{deq})}\label{computeLength}}
    \EndIf
    \If{$\var{num\sub{enq}} + \var{num\sub{deq}} = 0$}
        \State \Return \var{null}
    \Else
        \State \Return \var{new}
    \EndIf
\EndFunction{CreateBlock}

@@

\Function{<int, int>}{IndexDequeue}{\var{Node v, int b, int i}}
    \State \linecomment{return \var{<x, y>} such that \var{i}th dequeue in D(\var{v.blocks[b]})}
    \State \linecomment{is \var{y}th dequeue of D(\var{root.blocks[x]})}
    \State \linecomment{Precondition: \var{v.blocks[b]} has been propagated to root}
    \State \linecomment{and $\var{v.blocks[b].num\sub{deq}}\geq \var{i}$}
    \If{$\var{v} = \var{root}$}
        \State\Return \var{<b, i>} \label{indexBaseCase}
    \Else
	    \State \var{dir \assign\ (v.parent.left = v ? left : right)} 
    	\State \var{sup \assign\ v.blocks[b].super}
	    \If{$\var{b} > \var{v.parent.blocks[sup].end\sub{dir}}$}
	        \State{\var{sup} \assign\ $\var{sup}+1$}
	    \EndIf
	    \State \linecomment{compute index \var{i} of dequeue in superblock}
	    \State \hangbox{\var{i += v.blocks[b-1].sum\sub{deq}} - \\
	    		\var{v.blocks[v.parent.blocks[sup-1].end\sub{dir}].sum\sub{deq}}}
        \If{$\var{dir} = \var{right}$} 
        	\State \hangbox{\var{i += v.blocks[v.parent.blocks[sup].end\sub{left}].sum\sub{deq}} -\\
					\var{v.blocks[v.parent.blocks[sup-1].end\sub{left}].sum\sub{deq}}\label{considerLeftBeforeRight}}
        \EndIf \label{computeISuperEnd}
        \State \Return\Call{IndexDequeue}{v.parent, sup, i}
    \EndIf
\EndFunction{IndexDequeue}

@@

\Function{element}{FindResponse}{\var{int b, int i}}
    \State \linecomment{find response to \var{i}th dequeue in \var{root.blocks[b]}}
    \If{\var{root.blocks[b-1].\size + root.blocks[b].num\sub{enq}} $<$ \var{i}}\label{checkEmpty}
        \State \linecomment{queue is empty when the dequeue is linearized}
        \State \Return \var{null} \label{returnNull}
    \Else \ \linecomment{response is the \var{e}th enqueue in the root}
        \State $\var{e} \assign\ \var{i} + \var{root.blocks[b-1].sum\sub{enq}} - 
	\var{root.blocks[b-1].size}$\label{computeE}
		\State \linecomment{compute enqueue's block using binary search}
		\State find min $\var{b}'\leq \var{b}$ with $\var{root.blocks}[\var{b}'].\var{sum\sub{enq}} \geq \var{e}$
		\State \linecomment{find rank of enqueue within its block}
		\State $\var{i}' \assign\ \var{e} - \var{root.blocks}[\var{b}'-1].\var{sum\sub{enq}}$
        \State \Return \Call{GetEnqueue}{root, $\var{b}'$, $\var{i}'$}\label{findAnswer}
    \EndIf
\EndFunction{FindResponse}

@@

\Function{<int, int>}{DoublingSearch}{\sl{int} e, \sl{int} end}
\State \tt{start \assign\ end-1} \label{dsearchStart}
\While{\tt{root.blocks[start].sum\sub{enq}}$\geq$\tt{e}}
\State \tt{start \assign\ max(start-(end-start), 0)} \label{doubling}
\EndWhile \label{dsearchEnd}
\State \tt{b \assign\ root.BinarySearch(e, start, end)} \label{dsearchBinarySearch}
\State \tt{i \assign\ e- root.blocks[b-1].sum\sub{enq}} \label{DSearchComputei}
\State\Return \tt{<b,i>}
\EndFunction{DoublingSearch}
\end{algorithmic}
\end{algorithm}

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{blocks[b].num\sub{enq}$\geq$i$\geq 1$}}
\Function{element}{GetEnqueue}{\var{Node v, int b, int i}} \cmt{Returns the \tt{element} of $E_\tt{i}(\tt{this},\tt{b})$.}
\If{\tt{this} \bf{is} \tt{leaf}}
\State\Return \tt{blocks[b].element} \label{getBaseCase}
\ElsIf{\tt{i <= blocks[b].num\sub{enq-left}}} \label{leftOrRight} \cmt{$E_\tt{i}(\tt{this},\tt{b})$ is in the left child of this node.}
\State \tt{subblockIndex \assign left.BinarySearch(i+blocks[b-1].sum\sub{enq-left}, blocks[b-1].\eleft+1,} \label{leftChildGet}
\Statex \hspace{10.7em}\tt{blocks[b].\eleft)} 
\State \Return\tt{left.}\Call{GetEnqueue}{subblockIndex, i} 
\Else
\State \tt{i \assign\ i-blocks[b].num\sub{enq-left}}
\State\tt{subblockIndex \assign\ right.BinarySearch(i+blocks[b-1].sum\sub{enq-right}, blocks[b-1].\eright+1,} \label{rightChildGet}
\Statex \hspace{11.1em}\tt{blocks[b].\eright)} 
\State \Return\tt{right.}\Call{GetEnqueue}{subblockIndex, i} 
\EndIf
\EndFunction{GetEnqueue}
